Wide Post-Mass Transfer White-Dwarf Binaries

### 2024-07-30 summaries/paper-draft-plain.tex
add explanation for stable MT

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[top=0.60in, bottom=0.80in, left=0.60in, right=0.60in]{geometry}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{amsmath}
\usepackage[labelfont=bf]{caption}

\newcommand{\Msun}{\,\mathrm{M_{\odot}}}
\newcommand{\Rsun}{\,\mathrm{R_{\odot}}}
\newcommand{\MWD}{M_{\mathrm{WD}}}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\alphace}{\alpha_{\mathrm{CE}}}
\newcommand{\alphath}{\alpha_{\mathrm{th}}}
\newcommand{\alpharec}{\alpha_{\mathrm{rec}}}
\newcommand{\Ebind}{E_{\mathrm{bind}}}
\newcommand{\Egrav}{E_{\mathrm{grav}}}
\newcommand{\Eint}{E_{\mathrm{int}}}
\newcommand{\Eth}{E_{\mathrm{th}}}
\newcommand{\Erec}{E_{\mathrm{rec}}}
\newcommand{\Porb}{P_{\mathrm{orb}}}
\newcommand{\au}{\, \mathrm{au}}

@@

Secondly, a parameter on which stable mass transfer process depends is the accrection efficiency $\beta$, which in COSMIC is represented by \verb|acc_lim| in \verb|BSEDict|. Accretion efficiency is the fraction $\beta$ of all transfered materials that is accreted by the accretor. That is, $\dot{M_a} = -\beta\dot{M_d}$ where $M_d$ is the mass of donor and $M_a$ is the mass of accretor.

@@

Typical \textbf{pre-spike} and \textbf{spike} systems both follows the evolution path $\boxed{\text{\textbf{3-5-7-8-4-3-4}}}$ before reacing WD+MS. Note that the binaries come into contact (\verb|evol_type = 5|), which follows by a common enevelope phase. After inspecting the evolutionary process more closely, we found that the second stable mass transfer tends to widens the orbit significantly. \textbf{(***TO-DO***)}

>>>>

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[top=0.60in, bottom=0.80in, left=0.60in, right=0.60in]{geometry}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{physics}
\usepackage[labelfont=bf]{caption}

\newcommand{\Msun}{\,\mathrm{M_{\odot}}}
\newcommand{\Rsun}{\,\mathrm{R_{\odot}}}
\newcommand{\MWD}{M_{\mathrm{WD}}}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\alphace}{\alpha_{\mathrm{CE}}}
\newcommand{\alphath}{\alpha_{\mathrm{th}}}
\newcommand{\alpharec}{\alpha_{\mathrm{rec}}}
\newcommand{\Ebind}{E_{\mathrm{bind}}}
\newcommand{\Egrav}{E_{\mathrm{grav}}}
\newcommand{\Eint}{E_{\mathrm{int}}}
\newcommand{\Eth}{E_{\mathrm{th}}}
\newcommand{\Erec}{E_{\mathrm{rec}}}
\newcommand{\Porb}{P_{\mathrm{orb}}}
\newcommand{\au}{\, \mathrm{au}}
\newcommand{\Md}{M_{\mathrm{d}}}
\newcommand{\MT}{M_{\mathrm{T}}}
\newcommand{\Ma}{M_{\mathrm{a}}}

@@

Secondly, a parameter on which stable mass transfer process depends is the accrection efficiency $\beta$, which in COSMIC is represented by \verb|acc_lim| in \verb|BSEDict|. Accretion efficiency is the fraction $\beta$ of all transfered materials that is accreted by the accretor. That is, $\dot{\Ma} = -\beta\dot{\Md}$ where $\Md$ is the mass of donor and $\Ma$ is the mass of accretor.

@@

Typical \textbf{pre-spike} and \textbf{spike} systems both follows the evolution path $\boxed{\text{\textbf{3-5-7-8-4-3-4}}}$ before reacing WD+MS. Note that the binaries come into contact (\verb|evol_type = 5|), which follows by a common enevelope phase. After inspecting the evolutionary process more closely, we found that the second mass transfer (the stable one) tends to widens the orbit significantly.

To qualitatively explain the widening of the orbits, notice that for stable and non-conservative mass transfer with $\dot \Ma = - \beta \dot \Md$, we have
\[
  \frac{\dot J}{J} \propto (1 - \beta) q_d \frac{\dot \Md}{\MT},
\]
where $q_d = \Md / \Ma$ is the mass ratio between the donor and the accretor. Also,
\[
  \frac{\dot a}{a} = 2 \frac{\dot J}{J} - 2 \frac{\dot \Md}{\Md} - 2 \frac{\dot \Ma}{\Ma} + \frac{\dot \MT}{\MT}.
\]
Hence, for stable and non-conservative amss transfer, the widening of orbit can be evaluated with
\begin{equation}
  \frac{\dot a}{a} \propto - \qty(\frac{2}{q_d} - 2q_d - \beta) \frac{\dot \Md}{\MT}.
  \label{stableMT}
\end{equation}
Also note that the mass transfer begins when the primary is has \verb|evol_type = 8|, a naked Helium star on the Hertzsprung gap, whose wind mass loss is significant. For a fast wind mass loss, we have
\begin{equation}
  \frac{\dot a}{a} = - \frac{\dot \Md}{\MT}.
  \label{wind}
\end{equation}
This means that strong wind wtripping mass from the primary will cause the orbit to expand even further. To illustrate this, we change the \verb|windflag| in \verb|BSEDict| to 0, which correspond to weak wind. Now the $a_i = 2.80 \au$ system has final separation $a_f = 0.05 \au$, and the $a_i = 2.90 \au$ system has final separation $a_f = 0.14 \au$, both smaller than the original $0.09 \au$ and $0.27\au$ that correspond to stronger wind.

For our systems, the mass ratio $q_d > 1$, and the value of $q_d$ increases as the initial separation increases. Hence, based on Equation \ref{stableMT}, the stable mass transfer process actually shrinks the orbit, while the wind mass loss compensates it by windening the orbit. For \textbf{spike} bianries, although $q_d$ is slightly higher, causing the orbit to shrink more according to Equation \ref{stableMT}, the mass loss rate is significantly higher than \textbf{pre-spike} binaries, leading to a more significant increase in orbital separation. Also, \textbf{spike} binaries have larger orbital separation $a$ before mass transfer, which also leads to more significant increase in separation during the mass transfer.

For \textbf{merge} binaries, we notice that the separation of binaries at the start of CE is smaller than \textbf{spike} binaries. This lack of orbital energy to eject the CE explains the merger event. As initial separation contines to incerase, the binaries possess more energy to eject the CE after coming into contact, explaining the reappearance of WD+MS binaries instead of merger.

This concludes our qualitative analysis on the evolution results of $7\Msun + 1\Msun$ through stable mass transfer. Again, as shown in Figure \ref{stable_hi}, for stable mass transfer, the observed minimum separation of wide post-mass transfer WD+MS systems $0.15 \au$ can only be achieved in the \textbf{spike} region. That is, in our results, all wide post-mass transfer systems are due to the strong wind mass loss from the primary that widens the orbit significantly.
